We Are All Trying to Describe Something And We Don't Have the Words

### 2023-04-18 outputs/paper/paper.qmd
more lil edits
new mf formula. also bibliography

---
title: "We Are All Trying to Describe Something And We Don't Have the Words"
subtitle: "The Jack Antonoffification of Pop Music?"
author: 
  - Michaela Drouillard
thanks: "Code and data are available at: https://github.com/michaeladrouillard/jack."
date: today
abstract: "In this essay, I will be exploring Jack Antonoff's production style using data from the Spotify API. Drawing from an interview with Caleb Gamman, a guy who went viral for guessing which song on Taylor Swift's Midnights album was produced by Antonoff within seconds, I will delve into the specific characteristics that define an Antonoff-y song, and assess whether or not the audio features defined by Spotify can capture this. I plan to lasso Glen McDonald,the Spotify engineer behind these features, into explaining them to me. There will be a Bayesian inference model. It will be a logistic regression. We do not know whether or not it will be good. Finally, I will armchair a little bit about the cultural significance of Antonoff's music and the reasons behind his ubiquity in the music industry."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

Antonoff is prolific, to the extent that it is starting to bother people. He has produced and co-produced music for Lorde, Taylor Swift, Florence and the Machine, The Chicks, Clairo, The 1975, Grimes, Zayn, Pink, Lana Del Rey, Olivia Rodrigo, the Minions: The Rise of Gru Soundtrack, and more. All of this and his own bands too (Bleachers, and fun.). When I was getting brunch with a friend one Sunday and a Swift song came on, and she said "Jack Antonoff's gonna kill us all", we both laughed. What my friend was referring to, of course, is the fact that Jack Antonoff and his sound seems to be everywhere these days. As Andrew Marantz writes in the New Yorker: "When his band releases an album, the world responds politely. When he produces one by Lorde or Lana Del Rey or Taylor Swift, the world wobbles on its axis".

@@

Around the time that Midnights came out, I came across a video on Twitter of a guy, Caleb Gamman, guessing which songs on Midnights were produced by Antonoff within seconds. The tweet reads: "im able to instantly detect if jack antonoff worked on a song due to a visceral hatred of his production style". He nails it. Only one song on the album wasn't produced by Antonoff, and he guesses it.

I reached out to Gamman to ask him to explain Antonoff to me, only to find out, on the morning of the interview, that Gamman had released a 29 minute video essay on his Youtube Channel covering:

@@

I bolded the part about vocals sounding the way they would to the person who's singing them while editing this paragraph because I thought it was kind of funny, and maybe meaningful. It sounds like it's coming from inside the house!

@@

I pulled the data using the Spotify API, which I accessed using the `spotifyr` package in R. I used the `get_artist_audio_features` and `get_track_audio_analysis functions` function to acquire data on Lorde, Taylor Swift, St. Vincent, Lana Del Rey, The Chicks, Florence and the Machine, and Bleachers discographies. I threw in HAIM, Marina and the Diamonds, Maggie Rogers, Sharon Von Etten, and Mitski in too, since they've never collaborated with Jack Antonoff but are more or less part of the same pop space.

@@

I joined these data sets using the `rbind` function in base R, and created an additional variable called "jack", which contains a 1 if Jack Antonoff produced or co-produced the song, and a 0 if he had nothing to do with it. I got the information on which songs Antonoff did or did not produce from the "Jack Antonoff Production Discography" Wikipedia page \[cite\].

@@

The `get_artist_audio_features` provides a general profile of each song, according to variables developped by spotify, which makes for simple comparisons across artists. The final audio features in this study data set contains the variables: `artist_name`, `track_name`, `energy`, `danceability`, `key`, `loudness`, `mode`, `speechiness`, `acousticness`, `instrumentalness`, `liveness`, `valence`, `tempo`, `jack`.

@@

I thanked Glen for this time, ended the call, and booted up the ❤️statistical computing programming language, R ❤️\[citeR\].

#| label: fig-quaddanceability2
#| fig-cap: Danceability and Energy Quadrant
#| echo: false
#| warning: false

ggplot(data = df, aes(x = energy, y = danceability)) + 
  geom_point(aes(color = as.factor(jack)), alpha = 0.3) +
  geom_vline(xintercept = mean(df$energy), color = "gray") +
  geom_hline(yintercept = mean(df$danceability), color = "gray") +
  theme_minimal()

@@

ggplot(data = df, aes(x = energy, y = valence)) + 
  geom_point(aes(color = as.factor(jack)), alpha = 0.3) +
  geom_vline(xintercept = mean(df$energy), color = "gray") +
  geom_hline(yintercept = mean(df$valence), color = "gray") +
  theme_minimal()

@@

It looks like Jack-produced songs err towards danceability, across the spectrum of energy. It's a bit harder to gauge any trends from the valence quadrant. These quadrants are a start, but they don't suggest any defining Jack-iness characteristic underlying each of those blue points, which are scattered pretty broadly.

@@

#library(kableExtra)
#df |>
 # knitr::kable()

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

@@

$$
P(y=1|x) = \frac{1}{1 + e^{-(\beta_0 + \beta_1 x_1 + \cdots + \beta_n x_n)}}
$$

@@

After some minor pre-processing using the `tidyverse` and `lubridate` R packages. I followed Julia Silge's tutorial \[TK\] to build a model that deals with the imbalanced classes.

I used the `tidymodels` to fir the model. To deal with the imbalanced classes, I downsampled the majority class using the **`step_downsample()`** function from the **`themis`** package. Then I used 5-fold cross-validation with stratification by the **`jack`** variable to estimate the performance of the model.

@@

>>>>

---
title: "We Are All Trying to Describe Something And We Don't Have the Words"
subtitle: "The Jack Antonoffification of Pop Music?"
author: 
  - Michaela Drouillard
thanks: "Code and data are available at: https://github.com/michaeladrouillard/jack."
date: today
abstract: "In this essay, I will be exploring Jack Antonoff's production style using data from the Spotify API. Drawing from an interview with Caleb Gamman, a guy who went viral for guessing which song on Taylor Swift's Midnights album was produced by Antonoff within seconds, I will delve into the specific characteristics that define an Antonoff-y song, and assess whether or not the audio features defined by Spotify can capture this. I plan to lasso Glen McDonald,the Spotify engineer behind these features, into explaining them to me. There will be a Bayesian inference model. It will be a logistic regression. We do not know whether or not it will be good. Finally, I will armchair a little bit about the cultural significance of Antonoff's music and the reasons behind his ubiquity in the music industry."
format: pdf
number-sections: true
bibliography: bibliography.bib
---

@@

Antonoff is prolific, to the extent that it is starting to bother people. He has produced and co-produced music for Lorde, Taylor Swift, Florence and the Machine, The Chicks, Clairo, The 1975, Grimes, Zayn, Pink, Lana Del Rey, Olivia Rodrigo, the Minions: The Rise of Gru Soundtrack, and more. All of this and his own bands too (Bleachers, and fun.). When I was getting brunch with a friend one Sunday and a Swift song came on, and she said "Jack Antonoff's gonna kill us all", we both laughed. What my friend was referring to, of course, is the fact that Jack Antonoff and his sound seems to be everywhere these days. As Andrew Marantz writes in the New Yorker: "When his band releases an album, the world responds politely. When he produces one by Lorde or Lana Del Rey or Taylor Swift, the world wobbles on its axis" [@citeMarantz].

@@

Around the time that Midnights came out, I came across a video on Twitter of a guy, Caleb Gamman, guessing which songs on Midnights were produced by Antonoff within seconds [@citeTweet]. The tweet reads: "im able to instantly detect if jack antonoff worked on a song due to a visceral hatred of his production style". He nails it. Only one song on the album wasn't produced by Antonoff, and he guesses it.

I reached out to Gamman to ask him to explain Antonoff to me, only to find out, on the morning of the interview, that Gamman had released a 29 minute video essay on his Youtube Channel covering [@citeGamman]:

@@

He was really nice and fun to talk to. I bolded the part about vocals sounding the way they would to the person who's singing them while editing this paragraph because I thought it was kind of funny, and maybe meaningful. It sounds like it's coming from inside the house!

@@

I pulled the data using the Spotify API, which I accessed using the `spotifyr` package in R [@citeSpotify]. I used the `get_artist_audio_features` and `get_track_audio_analysis functions` function to acquire data on Lorde, Taylor Swift, St. Vincent, Lana Del Rey, The Chicks, Florence and the Machine, and Bleachers discographies. I threw in HAIM, Marina and the Diamonds, Maggie Rogers, Sharon Von Etten, and Mitski in too, since they've never collaborated with Jack Antonoff but are more or less part of the same pop space.

@@

I joined these data sets using the `rbind` function in base R, and created an additional variable called "jack", which contains a 1 if Jack Antonoff produced or co-produced the song, and a 0 if he had nothing to do with it. I got the information on which songs Antonoff did or did not produce from the "Jack Antonoff Production Discography" Wikipedia page [@citeWiki].

@@

The `get_artist_audio_features` provides a general profile of each song, according to variables developped by Spotify, which makes for simple comparisons across artists. The final audio features in this study data set contains the variables: `artist_name`, `track_name`, `energy`, `danceability`, `key`, `loudness`, `mode`, `speechiness`, `acousticness`, `instrumentalness`, `liveness`, `valence`, `tempo`, `jack`.

@@

I thanked Glen for this time, ended the call, and booted up the ❤️statistical computing programming language, R ❤️[@citeR].

```{r}
#| label: fig-quaddanceability2
#| fig-cap: Danceability and Energy Quadrant
#| echo: false
#| warning: false
library(ggplot2)
df |>
  ggplot(aes(x = energy, y = danceability)) + 
    geom_point(aes(color = as.factor(jack)), alpha = 0.3) +
    geom_vline(xintercept = mean(df$energy), color = "gray") +
    geom_hline(yintercept = mean(df$danceability), color = "gray") +
    theme_minimal()

@@

df |>
  ggplot(aes(x = energy, y = valence)) + 
    geom_point(aes(color = as.factor(jack)), alpha = 0.3) +
    geom_vline(xintercept = mean(df$energy), color = "gray") +
    geom_hline(yintercept = mean(df$valence), color = "gray") +
    theme_minimal()

@@

It looks like Jack-produced songs err towards danceability, across the spectrum of energy. It's a bit harder to gauge any trends from the valence quadrant, but it looks like Jack's songs tend to be somewhat on the side of low valence. These quadrants are a start, but they don't suggest any defining Jack-iness characteristic underlying each of those blue points, which are scattered pretty broadly.

@@

#library(kableExtra)
#df |>
 # knitr::kable()
```

@@

$$
\mbox{Pr}(y_i=1) = \mbox{logit}^{-1}\left(\beta_0+\beta_1 x_i\right)
$$

@@

After some minor pre-processing using the `tidyverse` R package [@citetidyverse]. I followed Julia Silge's tutorial [@citeSilge] to build a model that deals with the imbalanced classes.

I used the `tidymodels` to fit the model [@citetidymodels]. To deal with the imbalanced classes, I downsampled the majority class using the **`step_downsample()`** function from the **`themis`** package. Then I used 5-fold cross-validation with stratification by the **`jack`** variable to estimate the performance of the model.

@@

#even tho there's similrities, there's differences
